---
title: Interracial Straight Couples Meet
author: Angela Zhao 
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)

library(haven)

# Parameters
file_raw <- "/Users/angel/GitHub/dcl-2020-01/angela/c01-own/data/hcmst.rds"
#===============================================================================

# Code
```

## Data in its full glory

```{r}
#as_factor is the haven version

data_meet <-
  file_raw %>% 
  read_dta() %>% 
  mutate_if(is.labelled, as_factor) 

glimpse(data_meet)

```

# Interracial Couples, Straight


```{r}
# papglb_status: if respondent is gay, lesbian, or bisexual 

#398 total. 

data_meet %>% 
  filter(
    qflag != "no spouse or partner or otherwise unqualified", 
    respondent_race != partner_race, 
    same_sex_couple == "different sex couple"
  ) %>% 
  count(respondent_race, partner_race, sort = TRUE) 

```

## Met through Family or Friends

There doesn't seem to be many who have met through friend or family introductions. 

```{r}

data_meet %>% 
  filter(
    qflag != "no spouse or partner or otherwise unqualified", 
    respondent_race != partner_race, 
    same_sex_couple == "different sex couple", 
    q24_met_online == "met online"
  ) %>% 
  mutate(
    distinct_pair = 
      str_c(
        pmin(as.character(respondent_race), as.character(partner_race)), 
        "-", 
        pmax(as.character(respondent_race), as.character(partner_race))
      )
  ) %>% 
  select(distinct_pair, q24_R_cowork:q24_P_neighbor) %>% 
  pivot_longer(
    cols = q24_R_cowork:q24_P_neighbor, 
    values_to = "how_they_met", 
    names_to = "where_they_met"
  ) %>% 
  filter(how_they_met == "Yes") %>% 
  ggplot(aes(x = fct_infreq(where_they_met))) + 
  geom_bar()

```

## Met through Internet

38 met online, 348 met offline. 

More than half met as internet dating. The rest are split mostly evenly among internet games, chats, and social networking. Only one couple met through the internet community. 
  
```{r}

data_meet %>% 
  filter(
    qflag != "no spouse or partner or otherwise unqualified", 
    respondent_race != partner_race, 
    same_sex_couple == "different sex couple", 
    q24_met_online == "met online"
  ) %>% 
  mutate(
    distinct_pair = 
      str_c(
        pmin(as.character(respondent_race), as.character(partner_race)), 
        "-", 
        pmax(as.character(respondent_race), as.character(partner_race))
      )
  ) %>% 
  select(distinct_pair, q24_internet_dating:q24_internet_community) %>% 
  pivot_longer(
    cols = q24_internet_dating:q24_internet_community, 
    values_to = "how_they_met", 
    names_to = "where_they_met"
  ) %>% 
  filter(how_they_met == "Yes") %>% 
  ggplot(aes(x = fct_infreq(where_they_met))) + 
  geom_bar()

```
## Met through Public Places

About a quarter to a third met through a bar or restauarant, and then customer relationship and school. Then, it is mostly split between college, a volunteer organization, and church. Finally, military is the least common place to meet someone. 

```{r}

data_meet %>% 
  filter(
    qflag != "no spouse or partner or otherwise unqualified", 
    respondent_race != partner_race, 
    same_sex_couple == "different sex couple", 
  ) %>% 
  mutate(
    distinct_pair = 
      str_c(
        pmin(as.character(respondent_race), as.character(partner_race)), 
        "-", 
        pmax(as.character(respondent_race), as.character(partner_race))
      )
  ) %>% 
  select(distinct_pair, q24_school:q24_bar_restaurant) %>% 
  pivot_longer(
    cols = q24_school:q24_bar_restaurant, 
    values_to = "how_they_met", 
    names_to = "where_they_met"
  ) %>% 
  filter(how_they_met == "Yes") %>% 
  ggplot(aes(x = fct_infreq(where_they_met))) + 
  geom_bar()

```


